---
title: "Target Markdown"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(targets)
library(here)
library(dplyr)
library(ggplot2)
library(readr)
```

# Setup

Remove the `_targets_r` directory previously written by non-interactive runs of the report. Otherwise, the pipeline may contain superfluous targets.

```{r}
# library(targets)
# tar_unscript() # deletes the _targets.R
# tar_destroy("all") # removes parts or all of the objects from _targets/
```


# Targets

Targets are in `_targets.R` file.

Functions `tar_make()`, `tar_validate()`, `tar_manifest()`, `tar_glimpse()`, and `tar_visnetwork()` all need `_targets.R`. `_targets.R` lets these functions invoke the pipeline from a new external R process in order to ensure reproducibility.

```{r}
tar_validate() # Looks for errors.
```

```{r}
tar_manifest(fields = "command") # Data frame of target info.
```

```{r}
tar_glimpse() # Interactive dependency graph.
```

`tar_visnetwork()` also includes functions in the dependency graph, as well as color-coded status information.

```{r}
tar_visnetwork()
```

# Pipeline

If you ran all the `{targets}` chunks in non-interactive mode, then your R scripts are set up to run the pipeline.

```{r}
# tar_invalidate(names = run_plink_manual)
# tar_invalidate(names = cntrl_ids_folate_cov_file)
# tar_invalidate(names = starts_with("best_prs"))
# tar_invalidate(names = c(plink_prs_manual, moba_bim_updated))

tar_make()
```

# Output

You can retrieve results from the `_targets/` data store using `tar_read()` or `tar_load()`.

```{r, message = FALSE}
tar_objects()
tar_load(cntrl_ids_folate_cov)
cntrl_ids_folate_cov
```

```{r}
tar_load(gwas_summ_cleaned)
gwas_summ_cleaned
```

Genetic PCA

```{r}
tar_load(gen_pca)
gen_pca

gen_pca_eigenval <- read_delim(
  here("DATA", "cntrl_all_chr_QC_PCA.eigenval"),
  delim = "\t",
  col_names = "eigenval"
) %>%
  tibble::rowid_to_column("PC_nr")
gen_pca_eigenval

ggplot(gen_pca_eigenval) +
  geom_point(aes(as.factor(PC_nr), eigenval))
```

