---
title: "PRS - PRSice2"
author: "Julia Romanowska"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(targets)
library(readr)
library(dplyr)
library(ggplot2)
library(magrittr)
library(here)
library(patchwork)
library(ggiraph)
library(mgcv)
# library(xaringanthemer)
# style_solarized_light()

base_data <- "DATA/base_data_cleaned.txt"
target_data <- "DATA/cntrl_all_chr_QC"
pheno_data <- "DATA/mothers_IDs_folate_meas_blood.txt"
all_cov_data <- here("DATA", "covariates_4prs.txt")
out_prefix <- "RESULTS/MoBa_folat_prsice_with-cov_selected"
prs_levels <- 0.0001500
```

## Run script

The script was run in the console, using the following data / parameters.

```{r}
base_data
target_data
pheno_data
all_cov_data
out_prefix
prs_levels
```

This is the command used to run the script:

```{r, echo=FALSE}
sprintf(
    "Rscript PRSice.R --prsice PRSice_linux --base %s --target %s --binary-target F --pheno %s --base-maf MAF:0.01 --stat BETA --beta --cov %s --cov-factor 'diff_folate_intake' --bar-levels %s --fastscore --no-full --print-snp --out %s",
    base_data,
    target_data,
    pheno_data,
    all_cov_data,
    paste0(prs_levels, collapse = ","),
    out_prefix
  )
```

## Check results

Summary of the fit:

```{r}
summary <- read_delim(
  paste0(out_prefix, ".summary"),
  delim = "\t"
)
summary
```

Best fit:

```{r}
tar_load(best_prs_fit_manual)
best_prs_fit_manual

best_prs_fit_manual %>%
  count(In_Regression)
```

### Generating a quantile plot

Read the phenotype data and add to the PRS results.

```{r}
tar_load(cntrl_ids_folate_cov)
cntrl_ids_folate_cov

best_prs_fit_manual <- best_prs_fit_manual %>%
  select(-FID) %>%
  left_join(
    cntrl_ids_folate_cov,
    by = c("IID" = "SENTRIXID")
  )
best_prs_fit_manual %>%
  filter(is.na(SPFOLATE_F))

best_prs_fit_manual_all <- best_prs_fit_manual
best_prs_fit_manual <- best_prs_fit_manual %>%
  filter(!is.na(SPFOLATE_F))
```

The figure below shows how the measured values of folate concentration in blood
change with the quantiles of PRS.

```{r plot_prs_quant}
quant_plot <- ggplot(best_prs_fit_manual, aes(PRS_quant, SPFOLATE_F)) +
  geom_jitter() +
  # theme_xaringan() +
  xlab("deciles of PRS") +
  ylab("folate concentration in blood") +
  labs(
    title = "Grouping of women based on the decile of PRS\n and their folate concentration in blood"
  ) +
  theme(
    plot.title.position = "plot"
  )
smooth_plot <- ggplot(best_prs_fit_manual) +
  # geom_point(aes(PRS, SPFOLATE_F)) +
  geom_smooth(aes(PRS, SPFOLATE_F)) +
  # theme_xaringan()+
  xlab("PRS") +
  ylab("folate concentration in blood") +
  labs(
    title = "Smoothed PRS vs folate concentration",
    subtitle = "(using GAM method)"
  ) +
  theme(
    plot.subtitle = element_text(face = "italic"),
    plot.title.position = "plot"
  )

(quant_plot / smooth_plot) * theme_minimal() +
  plot_annotation(
    title = "PRSice2 - automated method",
    theme = theme(
      plot.title = element_text(face = "bold", size = 16, hjust = 0.5)
    )
  )

ggsave(
  here("RESULTS", "PRS_manual_fit.png"),
  width = 7,
  height = 5,
  units = "in",dpi = 120
)
```


```{r plot_prs_quant_article}
cor_test_prs <- correlation::cor_test(
  data = best_prs_fit_manual,
  x = "PRS",
  y = "SPFOLATE_F",
  method = "spearman"
)
cor_test_prs
plot(cor_test_prs, show_statistic = FALSE, show_ci = FALSE, stars = TRUE) +
  xlab("polygenic risk score (PRS)") +
  ylab("maternal plasma folate concentration (nmol/L)") +
  theme_minimal()

```


```{r gam_prs}
gam_model <- mgcv::gam(
  formula = SPFOLATE_F ~ s(PRS, bs = "cs"),
  data = best_prs_fit_manual,
  method = "REML"
)
gam_model
broom::tidy(gam_model)
```

Check residuals

```{r}
data_gam_fitted <- tibble(
  residuals = gam_model$residuals,
  fitted = gam_model$fitted.values
)
ggplot(data_gam_fitted, aes(fitted, residuals)) +
  geom_point()

ggplot(best_prs_fit_manual, aes(SPFOLATE_F)) +
  geom_histogram()

shapiro.test(best_prs_fit_manual$SPFOLATE_F)
shapiro.test(best_prs_fit_manual$orig_SPFOLATE_F)

```

### Check linearity using GAM

Only SPFOLATE_F ~ PRS

```{r gam}
broom::tidy(gam_model)

plot(gam_model)

gam.check(gam_model)
```

Adding covariates

```{r gam_cov}
tar_load(prs_covariate)
prs_covariate

best_fit_manual_cov <- best_prs_fit_manual %>%
  left_join(prs_covariate)
best_fit_manual_cov

gam_model_cov <- mgcv::gam(
  formula = SPFOLATE_F ~ s(PRS, bs = "cs",k = 5) +
    s(diff_folate_intake, bs = "cs", k = 6) +
    s(PC1, bs = "cs") +
    s(PC2, bs = "cs") +
    s(PC3, bs = "cs"),
  data = best_fit_manual_cov,
  method = "REML"
)
gam_model_cov
broom::tidy(gam_model_cov)

plot(gam_model_cov)
gam.check(gam_model_cov)
```

