---
title: "PRS - PLINK"
author: "Julia Romanowska"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(targets)
library(readr)
library(dplyr)
library(ggplot2)
library(magrittr)
library(here)
library(patchwork)
library(ggiraph)
library(mgcv)
library(xaringanthemer)
style_solarized_light()

base_data <- "DATA/extracted_gwas_summ_stats.txt"
target_data <- "DATA/cntrl_all_chr_QC"
pheno_data <- "DATA/mothers_IDs_folate_meas_blood.txt"
out_prefix <- "RESULTS/MoBa_folat_plink_selected"
all_cov_data_file <- here("DATA", "covariates_4prs.txt")
```

## The idea

Checking the PRS constructed from specific SNPs that were found in the
literature to affect the folate concentration. The SNPs proposed by Elisabeth:

```{r}
scan(here("DATA", "SNP_list_4test.txt"), what = "character")
```

However, some of those were not found in base data (GWAS summary stat) due to
e.g., quality issues or perhaps these were not included in the GWAS at all.

These SNPs were found in both, base data and MoBa data:

```{r}
read_table(base_data, progress = FALSE, col_types = NULL)
```

## Run script

The script was run in the console, using the following data / parameters.

```{r}
base_data
target_data
out_prefix
```

This is the command used to run the script:

```{r, echo=FALSE}
sprintf(
    "./plink --bfile %s --score %s 3 4 8 header --out %s",
    target_data,
    base_data,
    out_prefix
  )
```

Additionally, these data were used:

```{r}
pheno_data
all_cov_data_file
```

## Check results

Calculated score:

```{r}
tar_load(plink_prs_manual)
plink_prs_manual
```

Distribution of PRS and SPFOLATE_F:

```{r}
ggplot(plink_prs_manual, aes(SPFOLATE_F)) +
  geom_histogram()

ggplot(plink_prs_manual, aes(SCORE)) +
  geom_histogram()

```

### Generating a quantile plot

The figure below shows how the measured values of folate concentration in blood
change with the quantiles of PRS.

```{r plot_prs_quant}
quant_plot <- ggplot(plink_prs_manual, aes(PRS_quant, SPFOLATE_F)) +
  geom_jitter() +
  theme_xaringan() +
  xlab("deciles of PRS") +
  ylab("transformed folate conc.") +
  labs(
    title = "Grouping of women based on the decile of PRS\n and their folate concentration in blood",
    caption = "* folate concentration values transformed by inverse normal rank"
  ) +
  theme(
    plot.title.position = "plot"
  )
scatter_plot <- ggplot(plink_prs_manual) +
  geom_point(aes(SCORE, SPFOLATE_F)) +
  theme_xaringan()+
  xlab("PRS") +
  ylab("transformed folate conc.") +
  labs(
    title = "PRS vs folate concentration",
    caption = "* folate concentration values transformed by inverse normal rank"
  ) +
  theme(
    plot.subtitle = element_text(face = "italic"),
    plot.title.position = "plot"
  )

(quant_plot / scatter_plot) * theme_minimal() +
  plot_annotation(
    title = "PLINK - manual SNP selection",
    theme = theme(
      plot.title = element_text(face = "bold", size = 16, hjust = 0.5)
    )
  )

ggsave(
  here("RESULTS", "PRS_plink_manual_scatter.png"),
  width = 7,
  height = 5,
  units = "in",
  dpi = 120
)

smooth_plot <- ggplot(plink_prs_manual) +
  geom_smooth(aes(SCORE, SPFOLATE_F)) +
  geom_smooth(aes(SCORE, SPFOLATE_F), color = "navyblue", method = "lm") +
  theme_xaringan()+
  xlab("PRS") +
  ylab("transformed folate conc.") +
  labs(
    title = "PRS vs smoothed folate concentration",
    subtitle = "(using GAM and LM methods)",
    caption = "* folate concentration values transformed by inverse normal rank"
  ) +
  theme(
    plot.subtitle = element_text(face = "italic"),
    plot.title.position = "plot"
  )

(quant_plot / smooth_plot) * theme_minimal() +
  plot_annotation(
    title = "PLINK - manual SNP selection",
    theme = theme(
      plot.title = element_text(face = "bold", size = 16, hjust = 0.5)
    )
  )
ggsave(
  here("RESULTS", "PRS_plink_manual_smooth.png"),
  width = 7,
  height = 5,
  units = "in",
  dpi = 120
)

```


### Check linearity using GAM

Only SPFOLATE_F ~ PRS

```{r gam}
gam_model <- mgcv::gam(
  formula = SPFOLATE_F ~ s(SCORE, bs = "cs"),
  data = plink_prs_manual,
  method = "REML"
)
gam_model
broom::tidy(gam_model)

plot(gam_model)

gam.check(gam_model)
```

Adding covariates

```{r gam_cov}
tar_load(prs_covariate)
prs_covariate

plink_prs_manual_cov <- plink_prs_manual %>%
  left_join(prs_covariate)
plink_prs_manual_cov

gam_model_cov <- mgcv::gam(
  formula = SPFOLATE_F ~ s(SCORE, bs = "cs") +
    s(diff_folate_intake, bs = "cs", k = 6) +
    s(PC1, bs = "cs") +
    s(PC2, bs = "cs") +
    s(PC3, bs = "cs"),
  data = plink_prs_manual_cov,
  method = "REML"
)
gam_model_cov
broom::tidy(gam_model_cov)

plot(gam_model_cov)
gam.check(gam_model_cov)
```



