---
title: "Correlation between PRS and folate concentration"
subtitle: "stratified by type of ASM treatment"
author: "Julia Romanowska"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(tidyr)
library(dplyr)
library(purrr)
library(here)
library(stringr)
library(ggplot2)
library(naniar)
library(correlation)
library(see)
library(janitor)
library(flextable)
library(patchwork)
```

```{r read_data, message=FALSE}
source("00_params.R")

input_data <- read_delim(
  here("DATA", "cleaned_data.txt"),
  delim = "\t"
  ) %>%
  select(iid:NOAED_PREG_c, PLURAL, DUPLIKAT, AED_POLY, ends_with("_MONO"),
         FOL_PRE_or_TRIM1, FOLAT:autistic_trait_8yr) %>%
  pivot_longer(
    cols = starts_with("lang_imp"),
    names_to = "lang_imp_age",
    values_to = "lang_imp"
  ) %>%
  mutate(
    lang_imp_age = factor(
      stringr::word(
        lang_imp_age,
        start = -1,
        end = -1,
        sep = stringr::fixed("_")
      )
    )
  ) %>%
  mutate(MTHFR_2strat = ifelse(
    minor_al_load %in% c("CT", "TT"),
    yes = "CT/TT",
    no = "CC"
  )) %>%
  mutate(any_AED = ifelse(
    CC_STATUS == "epi_AED",
    yes = TRUE,
    no = FALSE
  )) %>%
  mutate(m_id = as.numeric(stringr::str_sub(M_ID_1111, 2, -1))) %>%
  mutate(
    across(
      c(minor_al_load, a1, a2, CC, IK_EPI, EPILEPSY2_c, FOL_PRE_or_TRIM1,
        starts_with("lang_imp"), starts_with("autistic"),
        AED_PREG_c, NOAED_PREG_c, CC_STATUS, MTHFR_2strat,
        AED_POLY, ends_with("_MONO")),
      ~ factor(.x)
    )
  )

all_ASM_vars <- unique(
  str_subset(
    names(input_data), pattern = "_MONO$"
  )
)
all_ASM_names <- all_ASM_vars %>%
  word(start = 1, end = 1, sep = fixed("_"))
names(all_ASM_vars) <- all_ASM_names

fol_suppl_diff <- read_delim(here("DATA", "folate_suppl_diff_covariate.txt"),
                             delim = "\t")

input_data <- input_data %>%
  left_join(fol_suppl_diff)
```

# Language impairment

```{r}
dependent <- "lang_imp"
dept_age_var <- "lang_imp_age"
cur_data <- input_data %>%
  select(iid, all_of(all_prs_types), CC_STATUS, EPILEPSY2_c, any_AED,
      AED_POLY:FOL_hydro, diff_folate_intake, starts_with("lang_imp"), m_id) %>%
  pivot_longer(
    cols = ends_with("_MONO"),
    names_to = "ASM_type",
    values_to = "ASM_type_use"
  )
all_ages <- levels(input_data$lang_imp_age)
```

## Frequency counts

```{r freq_count}
walk(all_ASM_vars, function(ASM){
  cat("\n       === CURRENT ASM TYPE: ", ASM, " ===\n\n")
  
  tmp_data <- cur_data %>%
    filter(
      !is.na(!!sym(dependent)) &
        ASM_type == ASM
    )
  print(
    janitor::tabyl(
      tmp_data, ASM_type_use, !!sym(dept_age_var)
    )
  )
})
```

The counts are really low - we will focus only on `CBZ_MONO`, `LTG_MONO`, `VPA_MONO`, and `LTA_MONO`.

```{r}
cur_ASM_vars <- all_ASM_vars[c("CBZ", "LTG", "VPA")]
```


```{r filter_asm_vars}
# NOTE: now, I am not interested in the age categories
cur_data <- cur_data %>%
  filter(ASM_type %in% cur_ASM_vars) %>%
  select(-!!sym(dept_age_var), -!!sym(dependent)) %>%
  distinct()
cur_data %>%
  count(ASM_type)
cur_data %>%
  mutate(has_fol_hydro = !is.na(FOL_hydro)) %>%
  janitor::tabyl(has_fol_hydro, ASM_type_use, ASM_type)
```

The tables above are counts for groups of those who used the ASM (`1`), those
who did not use (`0`), and those who have any data on folate concentration or not.

## Relation

### Stratified for ASM use

Plotting the relation between folate concentration and PRS stratified by use of
ASM, for the ASM types that we had most data for.

**Best-fit PRS**

```{r plot_relations_prs_auto}
ASM_labels <- c("carbamazepine", "lamotrigine", "valproate")
names(ASM_labels) <- as.character(cur_ASM_vars)
plot_corr_all_asm_prs_auto <- map(cur_ASM_vars, function(ASM){
    cur_title <- paste0(
      "ASM type: ", ASM_labels[ASM]
    )
    tmp_data <- cur_data %>%
      filter(ASM_type == as.character(ASM) & ASM_type_use == 1) %>%
      select(prs_auto, FOL_PRE_or_TRIM1, FOLAT, s_folat, diff_folate_intake,
             FOL_hydro)
    
    print(
      ggplot(
        tmp_data,
        aes(prs_auto, FOL_hydro)
        ) +
        geom_miss_point() +
        labs(
          title = cur_title,
          subtitle = "overview of missing and non-missing values"
        )
    )
    cur_plot <- tryCatch(
      {plot(
      cor_test(
        data = tmp_data,
        x = "prs_auto",
        y = "FOL_hydro",
        method = "spearman",
        partial = TRUE
        ),
        show_statistic = FALSE, show_ci = FALSE, stars = TRUE
      ) +
      labs(
        title = cur_title,
        caption = "Spearman correlation; adjusted for: FOL_PRE_or_TRIM1, FOLAT, s_folat, diff_folate_intake"
      ) +
      xlab("PRS (automatic) + covariates") +
      ylab("maternal plasma folate conc. (nmol/L)")},
      error = function(e){
        message("Could not perform adjusted spearman correlation! Plotting non-adjusted.")
        plot(
          cor_test(
          data = tmp_data,
          x = "prs_auto",
          y = "FOL_hydro",
          method = "spearman"
          ),
          show_statistic = FALSE, show_ci = FALSE, stars = TRUE
        ) +
        labs(
          title = cur_title,
          caption = "Spearman correlation; non-adjusted"
        ) +
        xlab("PRS (automatic)") +
        ylab("maternal plasma folate conc. (nmol/L)")
      }
    )
    print(cur_plot)
    return(cur_plot)
})
```

<br>
<br>

**2nd-best PRS**

```{r plot_relations_prs_manual}
ASM_labels <- c("carbamazepine", "lamotrigine", "valproate")
names(ASM_labels) <- as.character(cur_ASM_vars)
plot_corr_all_asm_prs_manual <- map(cur_ASM_vars, function(ASM){
    cur_title <- paste0(
      "ASM type: ", ASM_labels[ASM]
    )
    tmp_data <- cur_data %>%
      filter(ASM_type == as.character(ASM) & ASM_type_use == 1) %>%
      select(prs_manual, FOL_PRE_or_TRIM1, FOLAT, s_folat, diff_folate_intake,
             FOL_hydro)
    
    print(
      ggplot(
        tmp_data,
        aes(prs_manual, FOL_hydro)
        ) +
        geom_miss_point() +
        labs(
          title = cur_title,
          subtitle = "overview of missing and non-missing values"
        )
    )
    cur_plot <- tryCatch(
      {plot(
      cor_test(
        data = tmp_data,
        x = "prs_manual",
        y = "FOL_hydro",
        method = "spearman",
        partial = TRUE
        ),
        show_statistic = FALSE, show_ci = FALSE, stars = TRUE
      ) +
      labs(
        title = cur_title,
        caption = "Spearman correlation; adjusted for: FOL_PRE_or_TRIM1, FOLAT, s_folat, diff_folate_intake"
      ) +
      xlab("PRS (manual) + covariates") +
      ylab("maternal plasma folate conc. (nmol/L)")},
      error = function(e){
        message("Could not perform adjusted spearman correlation! Plotting non-adjusted.")
        plot(
          cor_test(
          data = tmp_data,
          x = "prs_manual",
          y = "FOL_hydro",
          method = "spearman"
          ),
          show_statistic = FALSE, show_ci = FALSE, stars = TRUE
        ) +
        labs(
          title = cur_title,
          caption = "Spearman correlation; non-adjusted"
        ) +
        xlab("PRS (manual)") +
        ylab("maternal plasma folate conc. (nmol/L)")
      }
    )
    print(cur_plot)
    return(cur_plot)
})
```


<br>
<br>

**rs1801133 genotype strata**

```{r plot_relations_MTHFR_strat}
t_tests_all_asm <- map(cur_ASM_vars, function(ASM){
    cur_title <- paste0(
      "ASM type: ", word(ASM, start = 1, end = 1, sep = fixed("_"))
    )
    tmp_data <- cur_data %>%
          filter(ASM_type == as.character(ASM) & ASM_type_use == 1) %>%
      select(MTHFR_2strat, FOL_PRE_or_TRIM1, FOLAT, s_folat, diff_folate_intake,
             FOL_hydro)
    
    print(
      ggplot(
        tmp_data,
        aes(MTHFR_2strat, FOL_hydro)
        ) +
        geom_miss_point() +
        labs(
          title = cur_title
        )
    )
    cat("T-test:\n")
    cur_t_test <- t.test(
      FOL_hydro ~ MTHFR_2strat,
      data = tmp_data
    )
    print(cur_t_test %>% broom::tidy())
    
    return(cur_t_test)
})
names(t_tests_all_asm) <- cur_ASM_vars

t_tests_all_asm
```

It doesn't look like there are any correlations here.

### Not stratitified

```{r plot_relations_all}
not_strat_prs_corr_plot <- c()
walk(all_prs_types, function(prs){
  tmp_data <- cur_data %>%
    filter(any_AED) %>%
    select(prs_manual, prs_auto, MTHFR_2strat, FOL_PRE_or_TRIM1, FOLAT, s_folat,
           diff_folate_intake, FOL_hydro) %>%
    distinct()
  
  print(
    ggplot(
      tmp_data,
      aes(!!sym(prs), FOL_hydro)
      ) +
      geom_miss_point() +
      labs(title = "overview of missing and non-missing values")
  )
  if(prs == "MTHFR_2strat"){
    cat("ANOVA test:\n")
    print(
    aov(
      formula = as.formula(
        paste0("FOL_hydro ~ ", prs,
               " + FOL_PRE_or_TRIM1 + FOLAT + s_folat + diff_folate_intake")
      ),
      data = tmp_data
    ) %>% broom::tidy()
    )
  } else if(prs == "prs_auto") {
    not_strat_prs_corr_plot <<- plot(
      cor_test(
        data = tmp_data %>%
          select(-MTHFR_2strat, -prs_manual),
        x = prs,
        y = "FOL_hydro",
        method = "spearman",
        partial = TRUE
        ),
        show_statistic = FALSE, show_ci = FALSE, stars = TRUE
      ) +
      labs(
        title = "any ASM use",
        caption = "Spearman correlation; adjusted for: FOL_PRE_or_TRIM1, FOLAT, s_folat, diff_folate_intake"
      ) +
      xlab("PRS (automatic) + covariates") +
      ylab("maternal plasma folate conc. (nmol/L)")
    print(not_strat_prs_corr_plot)
  } else {
    plot(
      cor_test(
        data = tmp_data %>%
          select(-MTHFR_2strat, -prs_auto),
        x = prs,
        y = "FOL_hydro",
        method = "spearman",
        partial = TRUE
        ),
        show_statistic = FALSE, show_ci = FALSE, stars = TRUE
      ) +
      labs(
        title = "any ASM use",
        caption = "Spearman correlation; adjusted for: FOL_PRE_or_TRIM1, FOLAT, s_folat, diff_folate_intake"
      ) +
      xlab("PRS (manual) + covariates") +
      ylab("maternal plasma folate conc. (nmol/L)")
  }
})

wrap_plots(
  c(plot_corr_all_asm_prs_auto, list(not_strat_prs_corr_plot)),
  ncol = 2) *
  theme_minimal() + 
  plot_annotation(tag_levels = "A", tag_suffix = ")")
ggsave(
  here("FIGURES", "corr_FOL_hydro_PRS_auto_strat_per_ASM_type.png"),
  height = 8,
  width = 10
)
```


## Counts per age

Just to be able to present something, we can count the cases/controls per age
category, ASM use, and MTHFR stratum. In the tables below, percentage of users
(`1`) and non-users (`0`) of each drug type is shown. The percentages are
calculated from totals per MTHFR-stratum (i.e., per column in the tables).

```{r}
freq_data <- input_data %>%
  select(iid, MTHFR_2strat, CC_STATUS, EPILEPSY2_c, any_AED,
         all_of(as.character(cur_ASM_vars)), starts_with("lang_imp"), m_id) %>%
  pivot_longer(
    cols = ends_with("_MONO"),
    names_to = "ASM_type",
    values_to = "ASM_type_use"
  )
```

```{r print_freq_detail}
counts_chosen_ASMs <- map(cur_ASM_vars, function(ASM){
  # cat("\n    ====== CURRENT ASM TYPE: ", ASM, " ======\n\n")
  counts_all_ages <- map(all_ages, function(cur_age){
    tmp_data <- freq_data %>%
      filter(
        !is.na(!!sym(dependent)) &
          ASM_type == as.character(ASM) &
          ASM_type_use == 1 &
          !!sym(dept_age_var) == cur_age
      )
    # cat("       === CURRENT AGE: ", cur_age, " ===\n")
      # print(
      out_table <- tabyl(
          tmp_data,
          !!sym(dependent),
          MTHFR_2strat
        ) %>%
          adorn_percentages("col") %>%
          adorn_pct_formatting(digits = 2) %>%
          adorn_ns() %>%
        as_tibble() %>%
        tibble::add_column(age = cur_age)
      # )
    # cat("\n")
      return(out_table)
  }) %>%
  bind_rows()
  return(counts_all_ages)
})
names(counts_chosen_ASMs) <- cur_ASM_vars

small_border <- officer::fp_border(color = "gray", width = 1)
header_border <- officer::fp_border(color = "gray30", width = 2)

all_counts_tables <- map(cur_ASM_vars, function(cur_ASM){
  cur_title <- paste("Women taking", cur_ASM)
  as_grouped_data(
    counts_chosen_ASMs[[cur_ASM]],
    groups = "age"
  ) %>%
    as_flextable() %>%
    bold(bold = TRUE, part = "header") %>%
    align(i = ~ !is.na(age), align = "center") %>% 
    bold(i = ~ !is.na(age)) %>%
    add_header_row(
      values = c("", "rs1801133 genotype"),
      colwidths = c(1, 2)
    ) %>%
    border_remove() %>%
    hline_bottom(
      j = 1:3,
      border = header_border
    ) %>%
    bg(
      i = ~!is.na(age),
      bg = "gray80"
    ) %>%
    border_inner_h(
      border = small_border,
      part = "body"
    ) %>%
    fix_border_issues() %>%
    width(
      j = c("CC", "CT/TT"),
      width = 4,
      unit = "cm"
    ) %>%
    set_caption(
      caption = cur_title
    )
})
names(all_counts_tables) <- cur_ASM_vars

all_counts_tables

walk(
  cur_ASM_vars,
  function(asm){
    cur_tab <- all_counts_tables[[asm]]
    save_as_docx(
      cur_tab,
      path = here(
        "RESULTS",
        paste0("counts_per_genotype_", dependent, "_asm_", asm, ".docx")
      )
    )
  }
)
```


# Autistic traits

```{r}
dependent <- "autistic_trait"
dept_age_var <- "autist_age"
# need to re-pivot data, now using the autistic_trait variables
cur_data <- input_data %>%
  select(iid, all_of(all_prs_types), CC_STATUS, EPILEPSY2_c, any_AED,
         AED_POLY:FOL_hydro, starts_with("autist"), m_id) %>%
  distinct() %>%
  pivot_longer(
    cols = starts_with("autistic"),
    names_to = dept_age_var,
    values_to = dependent
  ) %>%
  mutate(autist_age = stringr::str_sub(autist_age, -3, -1)) %>%
  pivot_longer(
    cols = ends_with("_MONO"),
    names_to = "ASM_type",
    values_to = "ASM_type_use"
  )
all_ages <- unique(cur_data[[dept_age_var]])
```

## Frequency counts

```{r freq_count}
```

The counts are really low for autistic traits as well. We might focus on the
same ASM types as above.

## Counts per age

Just to be able to present something, we can count the cases/controls per age
category, ASM use, and MTHFR stratum. In the tables below, percentage of users
(`1`) and non-users (`0`) of each drug type is shown. The percentages are
calculated from totals per MTHFR-stratum (i.e., per column in the tables).

```{r}
freq_data <- cur_data %>%
  select(iid, MTHFR_2strat, CC_STATUS, EPILEPSY2_c, any_AED,
         starts_with("ASM_"), starts_with("autist"), m_id) %>%
  filter(ASM_type %in% cur_ASM_vars)
```

```{r print_freq_detail}
```
